### 2019-12-11 defesa/tex/chapter/definicoes.tex
wip

An non-axis-parallel ellipse centered at $c\in\R^2$ can also be described by \autoref{def:ellipse}. Nonetheless, a simpler equation is going to be used here. Besides the center, the shape parameters $(a, b) \in \R^2_{>0}$, with $a > b$, an angle of rotation $\theta \in [0, \pi]$ is given. This angle represents the angle between the $x$-axis and the major-axis of the ellipse.

A rotated ellipse by $\theta$ can be transformed into a axis-parallel centered at the origin by applying two linear transformations. First, the ellipse can be translated to the origin and then rotated by $-\theta$, the reverse transformations are expressed by \autoref{eq:rotated_ellipse}.
\begin{equation}\label{eq:rotated_ellipse}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}=1.
\end{equation}
Also, the coverage region of an ellipse centered at $c\in\R^2$ with shape parameters $(a, b)\in\R^2_{>0}$ and rotation angle $\theta \in [0, \pi]$ is represented by every point $(x, y) \in \R^2$ that satisfies \autoref{eq:rotated_ellipse_co}.
\begin{equation}\label{eq:rotated_ellipse_co}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}\le 1.
\end{equation}

@@

A complex number is composed of a real part added to an imaginary part which is a multiple of $i = \sqrt{-1}$, this is expressed as $a + bi$, with $a, b \in \R$. Because $z$ is composed of two real numbers, the complex number system can be visualized on $\R^2$ as shown on \autoref{fig:complex-numbers}. This is also a good way to visualize Euler's Formula, as it can be also seen on \autoref{fig:complex-numbers}, which says that any complex number can be written in terms of its radius and polar angle as follows:
\begin{equation}
z = re^{i\theta},
\end{equation}
with $r$ being the length of the vector determined by the point $z$ on the complex plane and $\theta = angle(z)$ being its polar angle.
\begin{figure}[ht]
	\centering
	\def\svgwidth{\columnwidth}
	\includegraphics[scale=.5]{tex/figures/complex_numbers}
	\caption{The representation of a complex number on two dimensions.}
	\label{fig:complex-numbers}
\end{figure}

The complex conjugate is another important operator that is utilized later. Let $z = a + bi \in \mathbb{C}$, then we refer to $\bar{z}$ as the complex conjugate of $z$ and it is defined as:
\begin{equation}
\bar{z} = a - bi.
\end{equation}
Finally two observations about the angles of complex numbers are given here:
\begin{align}
angle(\bar{z}) = -angle(z)\\
angle(-z) = \pi - angle(z),
\end{align}
checking their validity can be done by just observing the symmetry between the points defined by $z, \bar{z},$ and $-z$ on the plane.

@@

In this work, only univariate polynomials are used, so the definition and results are focused on this case.

A function $p_n: \mathbb{C} \mapsto \mathbb{C}$ is a $n$-degree polynomial if it can be written as:
\begin{equation}\label{eq:poly}
p_n(z) = \sum_{k=0}^{n} a_k z^k.
\end{equation}
The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula\footnote{A finite number of $+, -, \times, \div, \sqrt{}$.} to determine their roots. However, a numerical approach exists for higher-degree polynomials.

@@

 Finding every eigenvalue of a matrix can be done using the QR algorithm, which runs in $\bigO(n^3)$ and uses $\bigO(n^2)$ memory. In \cite[p.~341]{golub} a practical version of it is developed. The idea is to convert the input matrix to the Hessenberg form, after that, the matrix preserves its form through the iterations and under some assumptions, convergence to the eigenvalues is achieved. It is worth mentioning that this approach is actually used in MATLAB and LAPACK, which is the external library used in NumPy for Python.
 Also, in \citeonline{barel} it is pointed that companion matrices are already in the Hessenberg form, making the Hessenberg conversion step unnecessary. This and other properties are used in \citeonline{barel} on the development of a $\bigO(n^2)$ algorithm for the specific case of companion matrices.

@@

The same definition found in \citeonline[p.~150]{powell} for real trigonometric polynomials is given here. They are also referred to as truncated Fourier Series in \citeonline{boyd:2006} and are given by:
\begin{equation}\label{eq:trig_poly}
p_n(\theta) = \sum_{k=0}^{n} a_k\cos(k\theta) + \sum_{k=1}^{n} b_k\sin(k\theta).
\end{equation}
We say that $p_n$ as defined by \autoref{eq:trig_poly} is a $n$-degree real trigonometric polynomial. An important property is stated in \citeonline[p.~150]{powell}, it says that a $n$-degree polynomial can have up to $2n$ distinct roots on the interval $[0, 2\pi)$. It also says that a function written in the format:
$$\begin{array}{cc}\cos^j{\theta}\sin^k{\theta} & \quad j, k \in \mathbb{Z}_+ \end{array},$$
can be transformed into a real trigonometric polynomial of degree $j+k$. Therefore, the following expression also represents a $n$-degree real trigonometric polynomial:
\begin{equation}\label{eq:trig_poly_2}
\sum_{0 \le j+k \le n}c_{j,k}\cos^j{\theta}\sin^k{\theta},
\end{equation}
for some $\{c_{j,k} \in \R: 0 \le j+k \le n\}$.

>>>>

A non-axis-parallel ellipse centered at $c\in\R^2$ can also be described by \autoref{def:ellipse}. Nonetheless, a simpler equation is going to be used here. Besides the center, the shape parameters $(a, b) \in \R^2_{>0}$, with $a > b$, an angle of rotation $\theta \in [0, \pi]$ is given. This angle represents the angle between the $x$-axis and the major-axis of the ellipse.

A rotated ellipse by $\theta$ can be transformed into a axis-parallel centered at the origin by applying two linear transformations. Reversing these transformations produces the following equation for a non-axis-parallel ellipse:
\begin{equation}\label{eq:rotated_ellipse}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}=1.
\end{equation}
Also, the coverage region of an ellipse centered at $c\in\R^2$ with shape parameters $(a, b)\in\R^2_{>0}$ and rotation angle $\theta \in [0, \pi)$ is represented by every point $(x, y) \in \R^2$ that satisfies \autoref{eq:rotated_ellipse_co}.
\begin{equation}\label{eq:rotated_ellipse_co}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}\le 1.
\end{equation}

@@

A complex number is composed of a real part added to an imaginary part which is a multiple of the imaginary unit $i = \sqrt{-1}$. This is expressed as $a + bi$, with $a, b \in \R$. Because $z$ is composed of two real numbers, the complex number system can be visualized on $\R^2$ as shown on \autoref{fig:complex-numbers}. This is also a good way to visualize Euler's Formula, as it can be also seen on \autoref{fig:complex-numbers}, which says that any complex number can be written in terms of its radius and polar angle as follows:
\begin{equation}
z = re^{i\theta}=r(\cos{\theta} + i\sin{\theta}),
\end{equation}
with $r$ being the length of the vector determined by the point $z$ on the complex plane and $\theta = angle(z)$ being its polar angle. Note that $angle$ is a function from $\mathbb{C}$ to $[0, 2\pi)$.
\begin{figure}[ht]
	\centering
	\def\svgwidth{\columnwidth}
	\includegraphics[scale=.4]{tex/figures/complex_numbers}
	\caption{The representation of a complex number on two dimensions.}
	\label{fig:complex-numbers}
\end{figure}

The complex conjugate is another important operator that is utilized later. Let $z = a + bi \in \mathbb{C}$, then we refer to $\bar{z}$ as the complex conjugate of $z$ and it is defined as:
\begin{equation}
\bar{z} = a - bi.
\end{equation}
Finally, we need to state two observations about the angles of $-z$ and $\bar{z}$ with respect to the angle of $z$, this will be useful later on \autoref{chapter:e3p}.
\begin{align}
angle(\bar{z}) = 2\pi-angle(z)\\
angle(-z) = \pi + angle(z),
\end{align}
checking the validity of these two equalities can be done by just observing the symmetry between the points defined by $z, \bar{z},$ and $-z$ on the plane.

@@

In this work, we are mostly interested in univariate polynomials defined over the complex numbers.

A function $p_n: \mathbb{C} \mapsto \mathbb{C}$ is a $n$-degree polynomial if it can be written as:
\begin{equation}\label{eq:poly}
p_n(z) = \sum_{k=0}^{n} a_k z^k,
\end{equation}
with $a_0, \dots, a_n \in \mathbb{C}$.
The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula\footnote{A finite number of $+, -, \times, \div, \sqrt{}$.} to determine their roots. Fortunately, a numerical approach exists for higher-degree polynomials which in practice works really well.

@@

 Finding every eigenvalue of a matrix can be done using the QR algorithm, which runs in $\bigO(n^3)$ and uses $\bigO(n^2)$ memory. A very complete introduction to this algorithm can be found on \citeonline{watkins:2008}. The idea is to convert the input matrix to the Hessenberg form, after that, the matrix preserves its form through the iterations and under some assumptions, convergence to the eigenvalues is achieved.
 Also, in \citeonline{barel} it is pointed that companion matrices are already in the Hessenberg form, making the Hessenberg conversion step unnecessary. This and other properties are used in \citeonline{barel} on the development of a $\bigO(n^2)$ algorithm for the specific case of companion matrices.

 In practice, LAPACK's ZGEEV routine is utilized--the user guide can be found in \citeonline{lapack}--which is an implementation of the QR algorithm that returns every eigenvalue of a complex matrix.

@@

The same definition found in \citeonline[p.~150]{powell} for real trigonometric polynomials is given here. They are also referred to as truncated Fourier Series in \citeonline{boyd:2006} and are given by:
\begin{equation}\label{eq:trig_poly}
p_n(\theta) = \sum_{k=0}^{n} a_k\cos(k\theta) + \sum_{k=1}^{n} b_k\sin(k\theta).
\end{equation}
We say that $p_n: \R \mapsto \R$ as defined by \autoref{eq:trig_poly} is a $n$-degree real trigonometric polynomial. An important property is stated in \citeonline[p.~150]{powell}, it says that a $n$-degree polynomial can have up to $2n$ distinct roots on the interval $[0, 2\pi)$. It also says that a function written in the format:
$$\begin{array}{cc}\cos^j{\theta}\sin^k{\theta} & \quad j, k \in \mathbb{Z}_+ \end{array},$$
can be transformed into a real trigonometric polynomial of degree $j+k$. Therefore, the following expression also represents a $n$-degree real trigonometric polynomial:
\begin{equation}\label{eq:trig_poly_2}
\sum_{0 \le j+k \le n}c_{j,k}\cos^j{\theta}\sin^k{\theta},
\end{equation}
for some $\{c_{j,k} \in \R: 0 \le j+k \le n\}$.

### 2019-12-11 defesa/tex/chapter/e3p.tex
wip

On \autoref{chapter:definitions} a brief introduction is given on how to get the roots of a polynomial. For that reason, we discuss two ways of converting $\xi$ into a polynomial. Symbolic computing was used to compute the polynomials, in practice an external Python library called SymPy was utilized (see \citeonline{sympy} for more details).

@@

It is possible to show that with that substitution and changing the variable to $z=e^{it}$, we obtain the following function $g : \mathbb{S} \mapsto \mathbb{C}$:

\begin{equation}
g(z)=\sum_{k=0}^{12} c_k z^{k-6},
\end{equation}
for some $c_0, \dots, c_{12} \in \mathbb{C}$. It is important to notice that this definition of $g$ does not impose that $z$ stays in the unit circle as any complex number, for that the additional condition: $|z|=1$ would need to be added.

Then, multiplying $g$ by $z^6 = e^{6i\theta}$, the following polynomial is obtained:

\begin{equation}
h(z) = z^6 g(z) = \sum_{k=0}^{12} c_k z^k.
\end{equation}

Then, every root of $\xi$ is a root of $h$. Note that the converse is not true, there could be roots of $h$ which are not roots of $\xi$, because a root $\hat{z}$ could have $|\hat{z}| \neq 1$ whereas $|e^{it}|$ is $1$.

>>>>

On \autoref{chapter:definitions} a brief introduction is given on how to get the roots of a polynomial. For that reason, we discuss two ways of converting $\xi$ into a polynomial in this section. Symbolic computing was used to compute the polynomials, in practice an external Python library called SymPy was utilized (see \citeonline{sympy} for more details).

@@

It is possible to show that with that substitution and changing the variable to $z=e^{i\theta}$, we obtain the following function $g : \mathbb{S} \mapsto \mathbb{C}$, with $\mathbb{S}$ being the unit circle ($\mathbb{S} = \{z \in \mathbb{C} : |z|=1\}$):

\begin{equation}
g(z)=\sum_{k=0}^{12} c_k z^{k-6},
\end{equation}
for some $c_0, \dots, c_{12} \in \mathbb{C}$. Even though $g$ is not a polynomial, it is easy to obtain one from it. Firstly, the negative exponentials need to be shifted, this can be done by just multiplying $g$ by $z^6$. Secondly, the domain cannot be restricted to the unit circle, so we define $h : \mathbb{C} \mapsto \mathbb{C}$ as:
\begin{equation}
h(z) = z^6 g(z) = \sum_{k=0}^{12} c_k z^k.
\end{equation}

Then, by its definition is possible to see that every root of $g$ is also a root of $h$. Conversely, every root $\hat{z}$ of $h$ which is in $\mathbb{S}$--or $|\hat{z}|=1$--is also a root of $g$. Naturally, the angle of a root $\hat{z}$ of $g$ is a root of $\xi$ by Euler's Formula.

It is possible to make another reduction to obtain a degree-$6$ polynomial from $h$ whose roots form a superset of the roots of $g$. As it has been mentioned on this chapter, an ellipse is symmetric with respect to its own axis. This means $\theta$ and $\pi + \theta$ are equivalent angles of rotation for any ellipse, thus $\xi(\theta) = \xi(\pi + \theta)$\footnote{$\xi$ is only defined over $[0, \pi)$, so this equality is not actually valid.}. 
On \autoref{chapter:definitions}, it was stated that the angle of $z$ and $-z$ has the same symmetry with each other as an ellipse's angle of rotation:
\begin{equation*}
angle(-z) = \pi + angle(z).
\end{equation*}
From that, as $g(e^{i\theta})=\xi(\theta)$ for every $\theta \in [0, 2\pi)$, we conclude that $g(-z)=g(z)$. This implies that $h$ is, in fact an even polynomial, or that $h(-z) = h(z)$ is true for every $z\in\mathbb{C}$:
\begin{align}
h(-z) = (-z)^6g(-z) = z^6g(z).
\end{align}
Therefore, all the odd degree coefficients of $h$ are $0$ and we can define the $6$-degree polynomial $f : \mathbb{C} \mapsto \mathbb{C}$ with the substitution $y=z^2$:
\begin{equation}
f(y) = \sum_{k=0}^{6} c_{2k} y^k.
\end{equation}
Then from every root $\hat{y}$ of $f$, two roots of $h$ can be obtained: $\sqrt{\hat{y}}$ and $-\sqrt{\hat{y}}$. As the angle of $-\sqrt{\hat{y}}$ is not between $[0, pi)$ we can ignore it. Note that the the square root of $\hat{y}$ does not need to be calculated, as only the angles are needed and they can be obtained by:
\begin{equation}
angle(\sqrt{z}) = angle(z)/2.
\end{equation}
Finally, using the QR algorithm mentioned on \autoref{chapter:definitions} a $\bigO(n^3)$ algorithm, with $n=6$, can be constructed for E3P.

It is also worth mentioning that a pattern on the coefficients of $f$ was identified, and maybe for future work it can be used for further improvements. Analyzing the polynomials produced for several instances, the following seems to true:
\begin{equation}
c_k = \hat{c_{6-k}},
\end{equation}
for $k=0, \dots, 6$. For now, we do not have any ideas on how it could be proved. Nevertheless, this seems to lead somewhere interesting because this condition guarantees that $f$ is a self-reciprocal polynomial which implies that its roots will always come in pairs $(\hat{z}, 1/\hat{z})$.
